noTCP: extending across middleboxes through unspecified behavior

### 2014-07-08 paper/paper.tex
first draft

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{noTCP: extend, extend, repurpose}
\ifnum\anon=1
\author{[Paper: \#XXX]}% \hspace{0.2cm} \ampmtime ]}
\else
\numberofauthors{2}
\author{
\alignauthor Andrius Aucinas\\
\affaddr{University of Cambridge} 
\and
\alignauthor Jon Crowcroft\\
\affaddr{University of Cambridge}
}
\fi
%for single author (just remove % characters)

% end author
\maketitle
% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
%\thispagestyle{empty}
\begin{abstract}
Too abstract.
\end{abstract}

@@

Extensibility of the Internet has been not like originally envisioned for a long time now, starting with IP layer and extending the \emph{narrow waist} to TCP or even HTTP at large. The problem primarily arises from the various middleboxes that perform packet content processing withing the network to improve performance\cite{Kopparty:2002ht,Chakravorty:2003dm}, increase security\cite{Handley:2001vp,Vutukuru:2008fc} and solve address shortage problems and is already well documented~\cite{UntoldMiddlebox2011,Qian:2012bj,Honda:2011ci,Guha:2005tb}.

The lack of extensibility has become acutely noticeable with the development of TCP extensions such as MPTCP and TCPCrypt, the former to make use of multi-homed endpoints, particularly useful in wireless networks and the latter to provide ubiquitous end-to-end traffic encryption. These and other extensions are primarily designed around the limitations of existing middleboxes, but they demonstrate two problems: 1) any extension design is extremely constrained by prevalent middlebox behaviors\cite{Honda:2011ci} and 2) we have already run out of TCP option space, the built-in mechanism for extending TCP.

Problems with the size of option space arise in MPTCP when its options are combined with those of regular TCP and therefore needs to change the semantics of a duplicate ACK~\cite{Handley:vj} and there are efforts to define an alternative control stream to work around the limitation of option space~\cite{Bonaventure:wx}. TCPCrypt goes even further and repurposes the Data portion of TCP segments to exchange cryptographic keys~\cite{Mazieres:uz}.

We propose \emph{noTCP}, an approach complementary to option negotiation which repurposes TCP header fields beyond what is specified but at the same time to be compatible with most of existing, deployed network infrastructure and safely fall-back to vanilla TCP if such modifications fail. At the core of the idea is modification of TCP header fields in a way that is not interfered with by middleboxes or so that such interference can be safely recovered or ignored.

@@

\begin{enumerate}
    \item Provide an analysis of current networks behavior in the presence of unspecified TCP protocol behavior.
    \item Propose a robust method to negotiate endpoint-specific behavior across middleboxes.
    \item Demonstrate that the method is effective across a number of cellular and WiFi networks.
\end{enumerate}

\section{Candidate protocol changes}

@@

\subsubsection*{Unset Flags}

The only header fields that may or may not have a meaning depending on the corresponding flag are \emph{urgent pointer} and \emph{acknowledgment number},with the latter only not being set at the initial SYN packet.

@@

The main difficulty with TCP Reserved bits is that there are only 3 bits unallocated if we consider the three ECN bits added over 10 years ago~\cite{Floyd:up,Ely:uc} but still not universally implemented~\cite{}. Any further reserved bit allocation would therefore be extremely difficult.

@@

Initial sequence numbers, must be hard to predict as they provide a measure (however weak) of security and until a stronger one (such as TCPCrypt) is universally adopted adding any information to the field in a way that is easily decoded by the receiver weakens security.

Window size field is potentially a better candidate, especially if used only during connection setup phase, since data is not typically transmitted during the handshake. We could further claim that even during data exchange a few of the least-significant bits could be sacrificed for carrying other information, however that would prevent flow control from operating properly. Higher-order bits could instead be given up when window scaling option is used to compensate for the decrease of signaled window size, but for every such bit the maximum window would be halved, having negative effects for more capable clients connected though high capacity links. Finally, we observed many middleboxes to modify the field and there is a risk of the middleboxes using the initially negotiated one for assigning resources for the duration of the connection~\cite{}.

There are multiple benefits of using the checksum field. Firstly, it is (only) used for error detection as opposed to other fields that govern protocol behavior and is either not always necessary due to the layered design of the OSI stack (when error detection is done at the layer below or above), or can be replaced with extensions' separate (MPTCP DSS checksum) or even stronger (TCPCrypt MAC) mechanisms. Secondly, checksum operations are efficient as they only involve one's complement addition. And finally, it is possible to generate checksums in a way that the intended value can recovered by the receiving end after traversing common types of middleboxes that recompute the checksum as we discuss in the following sections.

@@

The test has been implemented as an Android application and distributed through the \emph{Google Play Store} to collect data from volunteers. The overall structure of the tool is relatively straightforward:
\begin{itemize}
    \item UI which reports aggregated results to the end-users, sends results to the reporting server and manages the overall flow of the tests.
    \item Native implementation of the simplified TCP stack using Raw Unix sockets (\emph{SOCK\_RAW}) to allow us to generate packets at the network layer. This also meant having to drop RST packets generated by the kernel network stack during the test in response to incoming packets associated with connections it is not aware of - packets are delivered to both the raw socket and the kernel network stack.
    \item Similar server implementation that generates specific answers to incoming packets based on header and payload values.
\end{itemize}

Potential limitations of our methodology include that we do not consider packet resegmentation since the goal is to negotiate any endpoint-specific in packet headers so that correct behavior does not depend on how packets are split or coalesced together. To make sure that it does not happen all our packets contain minimal amount of payload (up to 16 bytes during data exchange) and data exchange only consists of a single message sent each way, in addition to any extra acknowlegements generated by middleboxes. Due to the same reason we also do not take into account congestion control.

\section{Network cahracteristics}
\label{sec:network}

\begin{table*}[t]
{\small
\begin{center}
\begin{tabular}{| l | >{\centering\arraybackslash}m{0.8cm} | >{\centering\arraybackslash}m{0.8cm} | >{\centering\arraybackslash}m{0.9cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
\hline
    \textbf{Net ID} & \textbf{Country} & \textbf{No filter} & \textbf{Global IP}   & \textbf{Port-specific} & \textbf{Validate checksum} & \textbf{Drop S/A data} & \textbf{Normalize ACK}  & \textbf{Normalize URG} & \textbf{Normalize Reserved} & \textbf{Remap Seq.} \\ \hline \hline
    % Univ Helsinki WiFi
    WiFi edu 1      & FI               &                    &                      &                        & \checkmark \checkmark      &                        &                         &                        &                             &                     \\ \hline
    % Eduroam
    WiFi edu 2      & UK               &                    & \checkmark           &                        &                            & \checkmark             &                         & \checkmark \checkmark  &                             & \checkmark          \\ \hline
    % wgb
    WiFi pub 1      & UK               & \checkmark         &                      &                        &                            &                        &                         &                        &                             &                     \\ \hline
    % Virgin home
    WiFi res 1      & UK               &                    &                      &                        & \checkmark \checkmark      &                        &                         &                        &                             &                     \\ \hline
    \hline
    % Finland internet.saunalahti
    Cellular 1      & FI               & \checkmark      &                      &                        &                            &                        &                         &                        &                             &                     \\ \hline
    % GiffGaff
    Cellular 2      & UK               &                 &                      &  \checkmark (443, 993) & -                          & -                      & -                       & -                      & -                           & -                   \\ \hline            
\end{tabular}
\end{center}
}
\caption{Network behaviour observed through tests generating custom TCP packets. A single checkmark means that particular behaviour was observed or in case of normalization the field was reset or replaced with a valid value. Double checkmark means that specific type of packet was discarded. A dash means that we observed different cases on the same network.}
\label{tab:networks}
\end{table*}

We show our collected results in Table~\ref{tab:networks}. Since we were focusing on mobile applications, we collected data across a number of public, residential and university WiFi networks and cellular networks in XX different countries: UK, USA, Finland (pending: Turkey, India, Lithuania, Spain...). We picked the most important dimensions that show what modifications to the protocol have the highest chance of being successull across the largest number of networks. In particular, the results include which networks provided clients a global IP address (very few), which ones we did not see any filtering for (a relatively large, but insufficient number to take any deployment for granted), and which ones discarded packets with invalid checksums, normalised particular fields - either dropped packets or reset to different values - and what type of mapping did NATs perform - only port number remapping or also changed sequence numbers.

Checksum behaviour was particularly interesting. In traditional NATs~\cite{Egevang:tu} checksum recomputation is very simple: subtract the old header fields from the checksum and add the new ones to minimize the resource use. In our sample X networks only do that, as indicated by the fact that they allow packets with incorrect checksums through towards the endpoint (otherwise the packet would either be dropped or a valid checksum would be recomputed). It is not always the case and another alternative is to add payload to a packet to make checksum a specific value. It can only be done when the receiving end is already expecting it, since it should not pass such payload up to the layer above, therefore can only be used from SYNACK onwards. A bigger issue here is that middleboxes do not always respect the standard and simply drop such packets even though the standard clearly allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}:

@@

We also tested middlebox itneraction with reserved header bits. We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe different behavior for these: all reserved bits were either allowed through or cleared at both phases. Importantly, we did not observe cases when such packets are dropped.

\subsection{Port-specific middlebox behavior}

One column of our table of observations is especially interesting -- port-specific behavior. We have only observed it on one network, but the particular network appears to be configured to allow traffic to commonly used SSL ports (443, 993) pass through unmodified. In fact, most of our attempted modifications are allowed through with the exception of packets that do not have a valid checksum being dropped. There may be other ports configured to allow unmodified traffic through, but we only focused on commonly used ports.

In addition to noting that our traffic was not encrypted but only sent to a specific destination port, it is also important to say that an increasing proportion of traffic is served over SSL. A commerciall deployed opt-in HTTP accelerator reports the proportion of HTTPS requests to be close to that of HTTP, although varying with time~\footnote{http://db.awazza.com/users/global/}.

All other traffic, however, is intercepted at the middlebox which we have separately confirmed to be a Bytemobile proxy~\footnote{Citrix Bytemobile - is it appropriate to reference this case accurately?} and it terminates clients' connections, establishing new ones to the destination, as well as adding options not present in original packets, changing window size, assigning new source port numbers sequentially and even postponing handshaking until the handshake with the client is complete.

\section{Protocol negotiation}

What is common to all candidate protocol changes described above is that they are very limited in size and therefore can only be used to exchange an opcode rather and not a substantial amount of information. Such information must instead be exchanged through other means - either implicit to the opcode, or explicit by redefining the meaning of part of the data payload or previously exchanged information. We therefore focus on combining the above observations into a robust exchange of the opcode between two endpoints in current networks.

>>>>

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{noTCP: extending across middleboxes through unspecified behavior}
\ifnum\anon=1
\author{[Paper: \#XXX]}% \hspace{0.2cm} \ampmtime ]}
\else
\numberofauthors{3}
\author{
\alignauthor Andrius Aucinas\\
\affaddr{University of Cambridge} 
\and
\alignauthor Jon Crowcroft\\
\affaddr{University of Cambridge}
\and
\alignauthor Narseo Vallina-Rodriguez\\
\affaddr{ICSI}
}
\fi
%for single author (just remove % characters)

% end author
\maketitle
% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
%\thispagestyle{empty}
\begin{abstract}
TCP has become the de facto transport protocol in the Internet. The necessity to extend it as well as the difficulty of doing so has lead researchers to explore new ways of exchanging control information between endpoints. In this paper we explore the choices for such exchange without using more TCP options. To achieve it we design noTCP - a mechanism which uses TCP header fields in combinations that do not have specified semantics to exchange a small amount of control information between endpoints to bootstrap a control channel. We also present a preliminary study which shows the mechanism works across large cellular and WiFi networks.
\end{abstract}

@@

Extensibility of the Internet has been not like originally envisioned for a long time now, starting with IP layer and extending the \emph{narrow waist} to TCP or even HTTP at large. The problem is caused by the various middleboxes that aim to improve performance\cite{Kopparty:2002ht,Chakravorty:2003dm}, increase security\cite{Handley:2001vp,Vutukuru:2008fc} and solve address shortage problems. Even congestion control has required multiple extensions since protocol's inception and the large number of proposed congestion control schemes~\cite{5462976} suggests that more will be necessary.

In an ideal world all infrastructure is forwards compatible and changes together with the needs of the users. Instead, TCP has become the de facto transport protocol of the Internet and development of its extensions has been very challenging. Proposals including MPTCP~\cite{Handley:vj} for multi-homed endpoints and TCPCrypt~\cite{Mazieres:uz} for ubiquitous end-to-end traffic encryption are necessary for today's networks but deployment has been slow.

Although TCP has been built to be extensible via options and reserved bits, all extensions are primarily designed around the limitations of existing network middleboxes\cite{Honda,Guha2005tb,UntoldMiddlebox2011}, but even if options were not filtered, we have run out of the 40-byte TCP option space.

Problems with the size of option space arise in MPTCP when its options are combined with those of regular TCP and therefore needs to change the semantics of a duplicate ACK~\cite{Handley:vj} and there are efforts to define an alternative control stream to work around the limitation of option space~\cite{Bonaventure:wx}. TCPCrypt uses the Data portion of TCP segments to exchange cryptographic keys~\cite{Mazieres:uz}. An alternative way of looking at the problem is to hide the information from the middleboxes as done by steganographic approaches~\cite{Frczek:2012dl,Zielinska:2014fn,Rowland:1997vq,Murdoch:2005fz}.

We propose \emph{noTCP}, an approach complementary to option negotiation which repurposes TCP header fields beyond what is specified. At the same time it is designed to work across most of deployed network infrastructure. At the core of the idea is modification of TCP header fields in a way that is not interfered with by middleboxes or so that such interference can be safely recovered or ignored. The idea builds up from the work on steganography using TCP~\cite{Frczek:2012dl,Murdoch:2005fz}, but we shift the focus from using sequence numbers and options to other fields as well.

@@

\begin{enumerate}
    \item Review the aspects of the TCP protocol that can potentially be used to extend it (Section~\ref{sec:protocolChanges}).
    \item Provide an analysis of current mobile (WiFi and cellular) network behavior in the presence of such modifications (Section~\ref{sec:network}).
    \item Propose a robust method to bootstrap control-stream between two endpoints across middleboxes (Section~\ref{sec:protocol}).
\end{enumerate}

\section{Candidate protocol changes}
\label{sec:protocolChanges}

@@

\subsubsection*{Unset Flags}
\label{sec:unset}

The only header fields that may or may not have a meaning depending on the corresponding flag are \emph{urgent pointer} and \emph{acknowledgment number}, and the latter is only allowed to be not set for the initial SYN packet.

@@

The main difficulty with TCP Reserved bits is that there are only 3 bits unallocated if we consider the three ECN bits added over 10 years ago~\cite{Floyd:up,Ely:uc} but still not universally implemented~\cite{Kuhlewind:2013hr}. Any further reserved bit allocation would therefore be extremely difficult.

@@

Initial sequence numbers must be hard to predict as they provide a measure (however weak) of security and until a stronger one (such as TCPCrypt) is universally adopted adding any information to the field in a way that is easily decoded by the receiver weakens security which has been problematic~\cite{Qian:2012bj,Bellovin:uz,Qian:2012wb}.

Window size field is potentially a better candidate, especially if used only during connection setup phase, since data is not typically transmitted during the handshake. We could further claim that even during data exchange a few of the least-significant bits could be sacrificed for carrying other information, however that would prevent flow control from operating properly. Higher-order bits could instead be given up when window scaling option is used to compensate for the decrease of signaled window size, but for every such bit the maximum window would be halved, having negative effects for more capable clients connected though high capacity links. Using the field risks interfering both with TCP flow control mechanism as well as window-size optimizations by the network~\cite{Kopparty:2002ht,Chakravorty:2003dm} performed \eg by Citrix ByteMobile Proxy~\cite{Ha:2006td}.

There are multiple benefits of using the checksum field. Firstly, it is only used for error detection as opposed to other fields that govern protocol behavior and is either not always necessary due to the layered design of the OSI stack (when error detection is done at the layer below or above), or can be replaced with extensions' separate (MPTCP DSS checksum) or even stronger (TCPCrypt MAC) mechanisms. Secondly, checksum operations are efficient as they only involve one's complement addition. And finally, it is possible to generate checksums in a way that the intended value can recovered by the receiving end after traversing common types of middleboxes that recompute the checksum as we discuss in the following sections.

@@

The test has been implemented as an Android application and distributed through the \emph{Google Play Store} to collect data from volunteers. The overall structure of the tool is relatively straightforward:
\begin{itemize}
    \item UI which reports aggregated results to the end-users, sends results to the reporting server and manages the overall flow of the tests.
    \item Native implementation of the simplified TCP stack using Raw Unix sockets (\emph{SOCK\_RAW}) to allow us to generate packets at the network layer. This also meant having to drop RST packets generated by the kernel network stack during the test in response to incoming packets associated with connections it is not aware of - packets are delivered to both the raw socket and the kernel network stack.
    \item Server implementation that generates specific answers to incoming packets based on header and payload values.
\end{itemize}

Potential limitations of our methodology include that we do not consider packet resegmentation since the goal is to negotiate any endpoint-specific in packet headers so that correct behavior does not depend on how packets are split or coalesced together. To make sure that it does not happen all our packets contain minimal amount of payload (up to 16 bytes during data exchange) and data exchange only consists of a single message sent each way, in addition to any extra acknowledgments generated by middleboxes. Due to the same reason we also do not take into account congestion control.

Instead of trying to infer which source port number would be assigned to a new connection by the kernel network stack without using it, we randomized the selection within the unprivileged port range. It was a simple optimization that works on any device and minimizes the probability of multiple tests failing due to another application using a particular port. We also repeated the tests on different destination ports, which we discuss in Sec.~\ref{sec:portspec}.

Non-transparent proxies are a major problem for extension deployment, therefore we used another well know and widely deployed tool to detect proxies on our tested networks. The tests were run separately and we used it to discover non-transparent proxies. This test works by checking if proxies respond to a TCP connection from the client: SYN request is going to a server that is instrumented not to reply to the client and the client would not be able to finish the handshake without a proxy, otherwise there is a proxy. To see if proxying depend on the higher-level protocol, it also detects HTTP proxies by checking content and header changes, different IP address seen by the server for HTTP or non-HTTP traffic, whether non-HTTP traffic on HTTP port is filtered, HTTP traffic is normalized (\ie deliberately modified headers are fixed) and some other aspects. Results are presented in Sec.~\ref{sec:proxies}.

\section{Network characteristics}
\label{sec:network}

\begin{table*}[t]
{\small
\begin{center}
\begin{tabular}{| l | >{\centering\arraybackslash}m{1.2cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
\hline
    \textbf{Net ID} & \textbf{Country} & \textbf{Port-specific} & \textbf{Validate checksum} & \textbf{Drop S/A data} & \textbf{Normalize Ack \#}  & \textbf{Normalize Urg ptr} & \textbf{Normalize Reserved} & \textbf{Remap Seq.} \\ \hline \hline
    % Univ Helsinki WiFi
    WiFi edu 1      & FI               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
    % Eduroam
    WiFi edu 2      & UK               &                        &                            & \checkmark             &                         & \checkmark             &                             & \checkmark          \\ \hline
    % wgb
    WiFi pub 1      & UK               &                        &                            &                        &                         &                        &                             &                     \\ \hline
    % MKSW germany (carlos)
    WiFi pub 2      & DE               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
    % Virgin home
    WiFi res 1      & UK               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
    % TEO Lt
    WiFi res 2      & LT               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
    % Telefonica Free
    WiFi ent 1      & ES               &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
    \hline
    % Finland internet.saunalahti
    Cellular 1      & FI               &                        &                            &                        &                         &                        &                             &                     \\ \hline
    % GiffGaff
    Cellular 2      & UK               &  443, 993              & -                          & -                      & -                       & -                      & -                           & -                   \\ \hline
    % E-Plus germany
    Cellular 3      & DE               &                        &                            &                        &                         & \checkmark             &                             & \checkmark          \\ \hline
    % EE/T-Mobile
    Cellular 4      & UK               &               80, 443  & -                          & -                      & -                       & -                      & -                           & -                   \\ \hline
    % Three UK
    Cellular 5      & UK               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
    % Omnitel LT
    Cellular 6      & LT               &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
\end{tabular}
\end{center}
}
\caption{Network behavior observed through tests generating custom TCP packets. A dash means that we observed different cases on the same network.}
\label{tab:networks}
\end{table*}

We show our collected results in Table~\ref{tab:networks}. Since we were focusing on mobile applications, in our preliminary study we collected results across a few public, residential, university and enterprise WiFi networks as well as 6 cellular networks in 4 different countries: UK, Finland, Germany and Lithuania.

The results include which networks did not filter any modifications (one WiFi and two cellular nets), and which ones discarded packets with invalid checksums, normalized particular fields - either dropped packets or reset to different values - and whether NATs performed only port number remapping or also changed sequence numbers. Only one network provided clients with a global IP (WiFi edu 2).

In traditional NATs~\cite{Egevang:tu} checksum recalculation is very simple: subtract the old header fields from the checksum and add the new ones to minimize the resource use. The networks that deliver packets with invalid checksum only do that as otherwise a packet with invalid checksum would be dropped.

A bigger issue we have discovered is that middleboxes do not always respect the standard and simply drop SYN-ACK packets with payload. We did not see it on cellular networks, but most WiFi (4/7) behaved in such way even though the standard allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}:

@@

As discussed in Section~\ref{sec:unset}, we can squeeze in a few extra bits of information by defining meaning to header fields when corresponding flags are not set. None of our tested networks filtered Acknowledgment Number field for the initial SYN packet on non-proxied paths and only two networks discarded packets with Urgent Pointer set without URG flag.

Finally we tested interaction reserved header bits. We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe different behavior for these: all reserved bits were either allowed through or cleared at both phases. Importantly, we did not observe cases when such packets are dropped.

\subsection{Port-specific middlebox behavior}
\label{sec:portspec}

We repeat our tests on a few ports:
\begin{itemize}
    \item 80, 443 - HTTP and HTTPS
    \item 993 - Secure IMAP port
    \item 5228 - Google cloud messaging port
    \item 6969 - random port number
    \item 8000 - common HTTP proxy port number
\end{itemize}

We have only observed port-specific behavior on two networks: \emph{Cellular 2} and \emph{Cellular 4}. In the first case, traffic travels over SSL ports (443, 993) unmodified except for packets with invalid checksum being dropped. Our traffic was not encrypted and an increasing proportion of traffic is served over SSL. A commercially deployed HTTP accelerator reports the proportion of HTTPS requests to be close to that of HTTP\footnote{\url{http://db.awazza.com/users/global/}}, making the value of such proxying questionable.

We confirmed the middlebox to be an optimizing proxy (ByteMobile). It replaces the original connections with a new one which adds options not present in original packets, changes window size and resets other fields to new values. Such behavior makes deployment of any protocol modifications or extensions very difficult (e.g. we verified that TCPCrypt does not work on this network) and although we did not see more cases of it, the company claims over 160 clients.

In the second case, traffic was only modified when going to ports 80 and 443. As we discuss below, the network optimizes HTTP traffic and since our traffic does not resemble valid HTTP, it is dropped.

\subsection{Non-transparent proxies}
\label{sec:proxies}

Non-transparent proxies are one of the biggest problems for TCP extension deployment. They terminate a connection between an initiator and destination and replaces it with a new one towards the destination, often deliberately changing connection parameters to optimize for certain behavior (\eg ByteMobile, HTTP proxies). We checked for non-transparent proxies on in our tested networks on a number of popular port numbers:

21, 22, 25, 80, 110,
  135,
  139,
  143,
  161,
  443,
  445,
  465,
  585,
  587,
  993,
  995,
 1194,
 1723,
 5060,
 6881,
 9001

\begin{table}[t!]
{\small
\begin{center}
\begin{tabular}{| l | >{\centering\arraybackslash}m{1.8cm} | >{\centering\arraybackslash}m{3.7cm} |}
\hline
    \textbf{Net ID} & \textbf{Radio Technology} & \textbf{Proxied Ports} \\ \hline \hline
    Cellular 2      & HSPA                      & 21, 22, 25, 80, 110, 135, 143, 161, 465, 585, 587, 995, 1194, 1723, 5060, 6881, 9001 \\ \hline
    Cellular 4      & GPRS                      & 80, 110, 143 \\ \hline
    Cellular 4      & HSPA                      & 80, 110, 143 \\ \hline
    Cellular 4      & LTE                       & 80, 110, 143, 443, 993, 995 \\ \hline
    % T-Mobile, part of EE
    Cellular 4\*     & HSPA                     & 25, 80, 110, 143 \\ \hline
    % Virgin on T-Mobile
    Cellular 4\*\*    & HSPA                    & 25, 80, 110, 143 \\ \hline
\end{tabular}
\end{center}
}
\caption{Proxies}
\label{tab:proxies}
\end{table}

Results are summarized in Table~\ref{tab:proxies}. Cellular 2 proxies almost all of our tested ports except for 443 and 993 as observed before. 
Cellular 4, is a more interesting case since reported proxying is not the same across the whole network. It is not obvious why ports used for encrypted communication are proxied over LTE and in our dataset there was only 1 case for HTTPS, but the other two cases seem related to the operator being formed by merging two operators. These differences are evident in \emph{whois} records, reported network name (\textbf{\*} shows the old operator name) and MVNO setup (\textbf{\*\*}). It is not safe to estimate proxying just from \emph{whois} records either: they do not necessarily match the operator names and we saw the same address ranges used with all Cellular 4 reported network names for HSPA and LTE radios.

Proxying is protocol-dependent in some cases: looking at port 80, Cellular 4\*\* modified HTTP headers, performed transcoding and reordered header fields, but only modified non-HTTP traffic for one of the two sessions. On Cellular 1 (15 sessions) two sessions show DNS proxies and on Cellular 5 (47 sessions) in two cases non-HTTP traffic is filtered on port 80, signaling that there still may be some proxying within the networks. Finally, Cellular 2 and 4 both reorder HTTP headers and perform transcoding, but allow non-HTTP traffic.

We did not have data about Cellular 6, but results discussed above suggest that there are no non-transparent proxies present. All other networks did not show non-transparent proxy behavior, but did not necessarily allow for TCP modifications either. In addition, the tests on port 80 show that a number of networks contain middleboxes that optimize HTTP, but such protocol-specific behavior is not uniform even within a single network making it more difficult to implement any changes.

%    an_op_name    | an_cell_net_type | an_country | port |    ip_global (simplified to network address. Included the owner according to whois
% db)
% -----------------+------------------+------------+------+-----------------
%  EE              | GPRS             | gb         |   80 | 31.72/19 (EE)
%  EE              | GPRS             | gb         |  110 | 31.72/19 (EE)
%  EE              | GPRS             | gb         |  143 | 31.72/19 (EE) 
%  EE              | HSPA             | gb         |   25 | 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE)
%  EE              | HSPA             | gb         |   80 | 31.80/21 (EE), 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE), 213.205.224/12 (EE)
%  EE              | HSPA             | gb         |  110 | 31.80/21 (EE), 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE), 213.205.224/12 (EE)
%  EE              | HSPA             | gb         |  143 | 31.80/21 (EE), 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE), 213.205.224/12 (EE)
%  EE              | LTE              | gb         |   80 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
%  EE              | LTE              | gb         |  110 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
%  EE              | LTE              | gb         |  143 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
%  EE              | LTE              | gb         |  443 | 213.205.224/12 (EE)
%  EE              | LTE              | gb         |  993 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
%  EE              | LTE              | gb         |  995 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
%  T-Mobile        | HSPA             | gb         |   25 | 31.96/21 (EE), 149.254/16 (T-Mobile)
%  T-Mobile        | HSPA             | gb         |   80 | 31.96/21 (EE), 149.254/16 (T-Mobile)
%  T-Mobile        | HSPA             | gb         |  110 | 31.96/21 (EE), 149.254/16 (T-Mobile)
%  T-Mobile        | HSPA             | gb         |  143 | 31.96/21 (EE), 149.254/16 (T-Mobile)
%  Virgin          | HSPA             | gb         |   25 | 149.254/16 (T-Mobile)
%  Virgin          | HSPA             | gb         |   80 | 149.254/16 (T-Mobile)
%  Virgin          | HSPA             | gb         |  110 | 149.254/16 (T-Mobile)
%  Virgin          | HSPA             | gb         |  143 | 149.254/16 (T-Mobile)
%  giffgaff        | HSPA             | gb         |   21 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |   22 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |   25 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |   80 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  110 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  135 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  143 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  161 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  465 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  585 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  587 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  995 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         | 1194 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         | 1723 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         | 5060 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         | 6881 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         | 9001 | 82.132/16 (O2)

\section{Protocol negotiation}
\label{sec:protocol}

What is common to all candidate protocol changes described above is that they are very limited in size and therefore sufficient to exchange an opcode or an ID. Other protocol details must instead be exchanged either out of band, implicit to the opcode, or explicitly by redefining the meaning of part of the data payload or previously exchanged information.

\subsection{Opcode embedding}

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figs/tcp-header}
% \vspace{-2mm}
\caption{TCP Header}
\label{fig:header}
% \vspace{-4mm}
\end{figure}

A surprising result in our tests is that the \emph{Acknowledgment Number} field without the ACK flag set is passed on all non-proxied connections. In addition, it is a 32-bit value and we can allocate part of it (\eg 8 or 16 bits) to signal that the rest of the field is significant. The exact allocation is subject to further discussion, but we suggest using the first 16 bits as \emph{Extension Space} and the others as $Opcode$ (Fig.~\ref{fig:header}), carrying the desired opcode. For simultaneous open we can use the same mechanism in both directions.

For acknowledging extensions in SYN-ACK packet we use Urgent Pointer and Checksum fields. Our tests indicate that neither choice is guaranteed to succeed, therefore we combine the two by redefining their values when SYN=1. We define Urgent Pointer to be $Opcode_U$ when URG=0 and Checksum to carry a meaning when Urgent Pointer = 0 to get around the cases where one is blocked but the other one is not. Across all networks we have tested in our preliminary study at least one of the choices succeeds.

Checksum value can not be used directly as an opcode since it changes at every level of NAT. The traditional NAT does a simple recalculation of the checksum~\cite{Egevang:tu}: subtracts the original source address and port number (in some cases also the sequence number) and adds the new values. The received Checksum after \(n\) NAT hops is therefore

\begin{align*}
Checksum_2 & = Checksum_1 - src_1 + src_2 \\
Checksum_i & = Checksum_{i-1} - src_{i-1} + src_i \\
 ... \\
Checksum_n & = Checksum_1 - src_1 + src_n
\end{align*}

Therefore we encode the Opcode as a target checksum so that it is recoverable by a simple binary subtraction of source port number and address and sequence numbers when they are remapped by the NAT (3-7 16-bit operations in total):

\begin{align*}
Checksum_{target} & = Opcode + src_1 \\
Checksum_n = Opcode_C & = Opcode + src_n
\end{align*}

There are two choices for setting checksum to a specific value: invalid checksum or adding padding payload. The latter can only be done when the receiving end is already expecting it, since it should not pass such payload up to the layer above, but keeps a packet valid from network's perspective.

For our prototype we assign the first 16 bits of the payload as \emph{Checksum Correction} to force a specific checksum, in line with other extensions which reuse part of data payload for option negotiation and maintaining validity of a packet. This Checksum Correction field is not delivered to the application. TLV-Encoded Control Stream can be present in data part~\cite{Bonaventure:wx} of the SYN-ACK packet after the Checksum Correction field.

Sending data with SYN-ACK packet works on all cellular networks we have tested, but does not on most WiFi. However, for all three cases (Urgent pointer and checksum) the information either reaches the destination as intended, or the packet is dropped. On all of our tested networks at least one case succeeds if there is no non-transparent proxy.

\subsection{Bootstrapping control stream}

We use the embedded opcodes in the handshake to exchange sufficient control information to bootstrap a complete control stream. We extend the three-way handshake with fields shown in Fig.~\ref{fig:header}:
